---
title: "What are some Factors that Prevent/Deter High School Students from Pursuing Higher Education?"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("analysis.R")
```

PowerNap by Iris Zhou
  \
  \
  
Blurb about percentage of college age students in college

Blurb about relation to high school graduation

The Washington Office of Superintendent of Public Instruction (OSPI) regularly publishes data about enrollment, demographics, finances, and graduation on their publicly accessible [data portal](https://www.k12.wa.us/data-reporting/data-portal) and in the [Washington Open Data Portal](https://data.wa.gov/). The OSPI published a [Report Card Graduation 2019-20](https://data.wa.gov/Education/Report-Card-Graduation-2019-20/gges-4vcv), which contains the most recent data enrollment, graduation rates, and dropout rates for students in 9th grade or higher. [Define cohort]


We used this data to look at the graduation rates across six student groups: foster care, non-foster care, homeless, non-homeless, low-income, and non-low income. 

```{r}
expanded_grad_rate_plot
```
  \
The graduation rates of foster care, homeless, and low-income students all fall below the overall graduation rate of `r round(all_students_grad_rate*100, 2)`%.


```{r student_group_outcomes_plot, fig.height=6, include=TRUE}
student_group_outcomes_plot
```

Although some students who did not graduate do continue to stay enrolled for the next school year, about a third of students in foster care and homeless students drop out. [How does this prevent pursuit of higher education]

```{r, fig.width=9, fig.height=5}
dropout_plots
```

As seen in the first graph, across all six of these student groups, students who end up dropping out tend to do so in the same timeline. The plurality of students who drop out tend to do so between their third and fourth year of high school. 

  
However, if we look at the proportion of students who drop out each year over the total number of students in that cohort, we see that the overall dropout rate sharply increases from Year 3 to 4 for homeless students and students in foster care. [drop out laws?]

  \
  \