---
title: "What are some Factors that Prevent/Deter High School Students from Pursuing Higher Education?"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("analysis.R")
```

PowerNap by Iris Zhou
  \
  \
  
According to the [National Center for Educational Statistics (NCES)](https://nces.ed.gov/programs/digest/d20/tables/dt20_302.60.asp), in 2019, only 40.7% of all 18-24 year olds were enrolled in college. The percentage of all 18-24 year olds enrolled in college has not changed significantly since 2008, when the enrollment rate was 39.6%. This begs the question, what are some factors preventing or deterring people from pursuing higher education?

One of the biggest stepping stones in the path toward pursuing higher education is the attainment of a high school diploma. Most colleges require either successful completion of high school, a GED certificate, or another degree equivalent for enrollment. Beyond the degree itself, many colleges also require students to have taken a certain number of courses in various subjects during high school.

The Washington Office of Superintendent of Public Instruction (OSPI) regularly publishes data about enrollment, demographics, finances, and graduation on their publicly accessible [data portal](https://www.k12.wa.us/data-reporting/data-portal). Their report for the 2019-2020 school year ([Report Card Graduation 2019-20](https://data.wa.gov/Education/Report-Card-Graduation-2019-20/gges-4vcv)) contains recent data on enrollment, graduation rates, and dropout rates for students in the ninth grade or higher.

I used this data to look at the graduation rates across six student groups: foster care, non-foster care, homeless, non-homeless, low-income, and non-low income.

```{r}
grad_rate_table
```


```{r}
grad_rate_plot
```
  \
The graduation rates of foster care, homeless, and low-income students, in contrast to the graduation rates of students not in those groups, all fall below the overall graduation rate of `r round(all_students_grad_rate*100, 2)`%. 


What happens to students who don't graduate? The data defines three possible outcomes for students who enter in 9th grade or who transfer in: graduating, dropping out, and continuing. 


```{r student_group_outcomes_plot, include=TRUE}
student_group_outcomes_plot
```

While some students who do not graduate do continue pursuing their secondary education, the majority of students who do not graduate drop out. More specifically, 38.8% of students in foster care, 31.6% of homeless students, and 18.7% of low-income students in Washington who enroll in high school end up dropping out. 

The OSPI also tracks the number of students in a cohort who dropout each year after enrolling in the 9th grade. They define a cohort as all the students who enroll in the 9th grade in an academic year, including students who transfer into this group, and excluding students who transfer out.

```{r cohort_dropout_rate_plot, include=TRUE}
cohort_dropout_rate_plot
```

Across all six of these student groups, students who end up dropping out tend to do so in the same timeline. The plurality of students who drop out tend to do so during their fourth year of high school. This is largely due to Washington State's mandatory attendance laws, which only allow children 16 or older to drop out if they are "regularly and lawfully employed", and either parental consent is given, or the child is emancipated ([RCW 28A.225.010](https://app.leg.wa.gov/rcw/default.aspx?cite=28a.225.010)). 

However, the dropout rate for students in the foster care, homeless, and low-income groups is consistently higher each year than for students not in those groups. The overall dropout rate increases much more sharply from Year 3 to 4 for homeless students and students in foster care. 


  \
  \